forrechelle

### 2019-12-20 analysis/20_forrechelle.Rmd
new figs

candidategenes <- c( "NR3C1", "NR3C2", "PRLR", "PRL", "POMC" ,"MC4R", "VIP", "BRCA1")
timepoints <- c("lay" , "inc.d9", "hatch", "n9")
hormones <- c("PRL", "CORT")

@@

```{r rechelle}

datapath <- "../results/"   # path to the data
datafiles <- dir(datapath, pattern = "*allvsd.csv") # get file names
datapathfile <- paste0(datapath, datafiles)

df <- datapathfile %>%
  setNames(nm = .) %>%
  map_df(~read_csv(.x, col_types = cols(), col_names = T), .id = "filename") %>% 
  mutate(tissue = sapply(strsplit(as.character(filename),'../results/06_'), "[", 2)) %>% 
  mutate(tissue = sapply(strsplit(as.character(tissue),'allvsd.csv'), "[", 1))  %>% 
  select(tissue, X1, everything()) 
head(df)

@@

df2$treatment <- factor(df2$treatment, levels = alllevels)

@@

  p <-  df2 %>%
    filter(tissue == i) %>%
    ggplot(aes(x = treatment, y = vsd, fill = gene, color = sex)) +
    geom_boxplot() +
    scale_alpha_manual(values = c(0.5, 1)) +
    labs(subtitle = i, y = "expression", x = "parental stage") +
    facet_wrap(~gene, scales = "free_y") +
    theme_B3() +
    scale_color_manual(values = sexcolors)
 print(p)
}

dft <- df %>% filter(X1 %in% candidategenes) %>%
  unite("tissuegene", tissue:X1, remove = FALSE) %>%
  select(-tissue, -`X1`, -filename)  %>%
  pivot_longer(-tissuegene, names_to = "samples", values_to = "vsd") %>%
  drop_na() %>%
  mutate("bird" = sapply(strsplit(as.character(samples),'\\_'), "[", 1)) %>%
  select(bird,tissuegene,vsd) %>%
  pivot_wider(id_cols = bird, names_from = tissuegene, values_from = vsd)

dft <- as.data.frame(dft)
row.names(dft) <- dft$bird
dft$bird <- NULL

@@

x %>% network_plot.cor_df(min_cor = .2)

x %>%
  focus(pit_PRL) %>%
  mutate(rowname = reorder(rowname, pit_PRL)) %>%
  ggplot(aes(rowname, pit_PRL)) +
    geom_col() + coord_flip() +
  labs(y = "corrleation with pituitiary PRL", x = "candidate genes by tissue") +
  theme_B3() + 
  theme(legend.position = "none")
```

>>>>

candidategenes <- c("PRLR", "PRL", "CISH", "JAK2", "SOCS3", "GH" )
timepoints <- charlevels

@@

```{r data}
datapath <- "../results/"   # path to the data
datafiles <- dir(datapath, pattern = "*allvsd.csv") # get file names
datapathfile <- paste0(datapath, datafiles)

df <- datapathfile %>%
  setNames(nm = .) %>%
  map_df(~read_csv(.x, col_types = cols(), col_names = T), .id = "filename") %>% 
  mutate(tissue = sapply(strsplit(as.character(filename),'../results/06_'), "[", 2)) %>% 
  mutate(tissue = sapply(strsplit(as.character(tissue),'allvsd.csv'), "[", 1))  %>% 
  select(tissue, X1, everything()) 
```

```{r forrechelle}
# check which genese appear in multiple tissue
tissuespecificity <- df %>%
  group_by(X1) %>%
  summarize(tissues = str_c(tissue, collapse = ", ")) %>%
  filter(tissues != "gon, hyp, pit") %>%
  group_by(tissues) %>%
  summarize(genes = str_c(X1, collapse = ", ")) %>% 
  mutate(n = str_count(genes, ",") + 1) %>% 
  arrange(desc(n))
tissuespecificity
write.csv(tissuespecificity, "../results/tissuespecificity.csv")

@@

df2$treatment <- factor(df2$treatment, levels = alllevels)
df2$tissue <- factor(df2$tissue, levels = tissuelevels)

@@

df3 %>%
    ggplot(aes(x = treatment, y = m)) +
    geom_errorbar(aes(ymin=m-se, ymax=m+se, color=gene), width=.1) +
    geom_point(size = 1, aes(color = gene)) +
    geom_line(aes(x = as.numeric(treatment), y = m, color = gene)) +
    scale_alpha_manual(values = c(0.5, 1)) +
    labs( y = "average expression", x = "parental stage") +
    facet_grid(sex ~ tissue, scales = "free_y") +
    theme_B3() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

df2$gene <- factor(df2$gene)

for (i in levels(df2$gene)) {
  p <-  df2 %>%
    filter(gene %in% i) %>%
    ggplot(aes(x = treatment, y = vsd, color = sex, fill = treatment)) +
    geom_boxplot() +
    scale_alpha_manual(values = c(0.5, 1)) +
    labs(y = "expression", x = "parental stage", subtitle = i) +
    facet_wrap(~ tissue , scales = "free_y", nrow = 3) +
    theme_B3() +
    scale_color_manual(values = sexcolors) +
    scale_fill_manual(values = colorscharmaip)
 print(p)
}

dft <- df %>% filter(X1 %in% candidategenes) %>%
  unite("tissuegene", tissue:X1, remove = FALSE) %>%
  select(-tissue, -`X1`, -filename)  %>%
  pivot_longer(-tissuegene, names_to = "samples", values_to = "vsd") %>%
  drop_na() %>%
  mutate("bird" = sapply(strsplit(as.character(samples),'\\_'), "[", 1)) %>%
  select(bird,tissuegene,vsd) %>%
  filter(bird != "x.g.g.ATLAS") %>%
  pivot_wider(id_cols = bird, names_from = tissuegene, values_from = vsd)
dft <- as.data.frame(dft)
row.names(dft) <- dft$bird
dft$bird <- NULL
head(dft)

@@

x %>% network_plot.cor_df() + 
  scale_color_gradient2(low = "blue", mid = "grey", high = "red", midpoint = 0, limits = c(-1,1))

a <- x %>%
  focus(hyp_PRL) %>%
  mutate(rowname = reorder(rowname, hyp_PRL)) %>%
  ggplot(aes(rowname, hyp_PRL, fill = hyp_PRL)) +
    geom_col() + coord_flip() +
  labs(y = "corr. w/ hyp PRL", x = "candidate genes by tissue") +
  theme_B3() + 
  theme(legend.position = "bottom", legend.title = element_blank()) +
  scale_fill_gradient2(low = "blue", mid = "grey", high = "red", midpoint = 0, limits = c(-1,1)) +
  ylim(c(-0.1,0.6))

b <- x %>%
  focus(pit_PRL) %>%
  mutate(rowname = reorder(rowname, pit_PRL)) %>%
  ggplot(aes(rowname, pit_PRL, fill = pit_PRL)) +
    geom_col() + coord_flip() +
  labs(y = "corr. w/ pit PRL", x = NULL) +
  theme_B3() + 
  theme(legend.position = "bottom", legend.title = element_blank()) +
  scale_fill_gradient2(low = "blue", mid = "grey", high = "red", midpoint = 0, limits = c(-1,1))  +
  ylim(c(-0.1,0.6))

c <- x %>%
  focus(gon_PRL) %>%
  mutate(rowname = reorder(rowname, gon_PRL)) %>%
  ggplot(aes(rowname, gon_PRL, fill = gon_PRL)) +
    geom_col() + coord_flip() +
  labs(y = "corr. w/ gon PRL", x = NULL) +
  theme_B3() + 
  theme(legend.position = "bottom", legend.title = element_blank()) +
  scale_fill_gradient2(low = "blue", mid = "grey", high = "red", midpoint = 0, limits = c(-1,1))  +
  ylim(c(-0.1,0.6))

plot_grid(a,b, c, nrow = 1, rel_widths = c(1.1,1,1))
```
